The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-03 analysis/paper/paper.qmd
Update description and comments in topic model code chunk

---
title: "The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach"
author:
  - Claudiu Forgaci:
      correspondence: "yes"
      email: C.Forgaci@tudelft.nl
      orcid: 0000-0003-3218-5102
      institute: tud
institute:
  - tud:
      name: Delft University of Technology
      address: Julianalaan 134, 2628 BL, Delft, Zuid-Holland, The Netherlands
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  urbanism; education; post-master; text mining
highlights: |
  These are the highlights. 
---

@@

```{r setup}
#| echo: false
#| warning: false

@@

```{r get-pdf-text}
#| echo: false
#| cache: true

data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text_and_location.csv")

if (!file.exists(data_path)) {
  # On the first run, this function might take extra time as a Python environment needs to be set up
  emu_theses$text <- convert_pdf_text(pdf_paths)

  # Write thesis dataframe to file
  write_csv(emu_theses, data_path)
} else {
  emu_theses$text <- read_csv(data_path)$text
  emu_theses$longitude <- read_csv(data_path)$longitude
  emu_theses$latitude <- read_csv(data_path)$latitude
}

@@

```{r clean-text}
#| echo: false

@@

In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program. To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.

@@

```{r fig-map-theses}
#| warning: false 
#| echo: false
#| fig-cap: "Location of EMU theses included in the analysis"

@@

@fig-map-theses shows the geographic spread of the theses included in the analysis.

@@

@@

```{r stop-words}

emu_theses$text_clean <- emu_theses$text_clean |>
  str_remove_all(find_meta_stopwords(emu_theses)) |> 
  str_remove_all(urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism", "tu delft", "ku leuven", "upcs barcelona", "iuav venice"))) |>
  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))

@@

Additional checks were made to the document in search for previously undetected stop words. The checks included identifying words consisting of four or less characters that extensively used in the text corpus.

```{r}
#| label: short-stop-words

#short_words(emu_theses$text_clean, scope = 'corpus')
short_words(emu_theses_words$word)

@@

```{r stemming}

@@

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017].

```{r}
# The {topicmodels} package requires a document-term matrix as input
emu_theses_dtm <- convert_to_dtm(emu_theses_words, "title", "word")
emu_theses_dtm

# We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling
# In LDA, each document is a mixture of topics and each topic is a mixture of words
emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
emu_theses_lda

@@

# The top 10 words that are the most common in each topic
emu_theses_top_terms <- emu_theses_topics |> 
  group_by(topic) |> 
  top_n(20, beta) |> 
  ungroup() |> 
  arrange(topic, -beta)

emu_theses_top_terms |> 
  mutate(term = reorder(term, beta)) |> 
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

@@

```{r}
# Which are the top n most frequently used words in each thesis?
get_top_words_per_document(data = emu_theses_words, 
                           top_n = 3, 
                           title_col = "title", 
                           word_col = "word")
```

```{r}
# Which are the top n most frequently used words in all theses`?
get_top_words_per_corpus(data = emu_theses_words, 
                         top_n = 5, 
                         word_col = "word")
```

```{r}
# Which are the top n most frequently used words in all theses`?
emu_theses_top_20 <- get_top_words_per_corpus(data = emu_theses_words, 
                         top_n = 20, 
                         word_col = "word")
```

>>>>

---
title: "The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute: tud
  - John Q. Doe:
      email: johndoe@fosg.org
      orcid: 0000-0003-1689-0558
      institute: tud
institute:
  - tud:
      name: Delft University of Technology
      address: Julianalaan 134, 2628 BL, Delft, Zuid-Holland, The Netherlands
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  urbanism; education; post-master; text mining
highlights: |
  These are the highlights. 
---

@@

```{r}
#| label: setup
#| echo: false
#| warning: false

@@

```{r}
#| label: extract-pdf-text
#| echo: false
#| cache: true

data_path <- here("analysis", "data", "derived_data", "emu_theses_with_text.csv")

if (!file.exists(data_path)) {
  # On the first run, this function might take extra time as a Python environment needs to be set up
  emu_theses$text <- convert_pdf_text(pdf_paths)
  write_csv(emu_theses, data_path)
} else {
  emu_theses <- read_csv(data_path)
}

@@

```{r}
#| label: clean-text
#| echo: false

@@

In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved through the years of the program. To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu_theses$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`. The first year of the program was not represented, as PDF files were only available from 2008 onward.

@@

```{r}
#| label: fig-map-theses
#| warning: false 
#| echo: false
#| fig-cap: "Location of EMU theses included in the analysis"

@@

Figure @fig-map-theses shows the geographic spread of the theses included in the analysis.

@@

[@silge2017a] [@hvitfeldt2021a]

@@

```{r}
#| label: stop-words

emu_theses$text_clean <- emu_theses$text_clean |>
  str_remove_all(find_meta_stopwords(emu_theses)) |> 
  str_remove_all(urbanism_stopwords(
    add_stopwords = c("emu", "european postgraduate master in urbanism"))) |>
  str_remove_all(thesis_stopwords(add_stopwords = c("advisor", "prof")))

@@

@@

```{r}
#| label: stemming

@@

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017]. We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. In LDA, each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the beta statistic. The probability of a document being part of a topic is given by the gamma statistic.

```{r}
# The topicmodels package requires a document-term matrix as input
emu_theses_dtm <- emu_theses_words |> 
  count(title, word, sort = TRUE) |> 
  cast_dtm(title, word, n)

emu_theses_dtm

# Fit LDA model
emu_theses_lda <- LDA(emu_theses_dtm, k = 2, control = list(seed = 2023))
emu_theses_lda

@@

# The top n words that are the most common in each topic
emu_theses_top_terms <- get_top_words_per_topic(emu_theses_topics, 20)

emu_theses_top_terms |> 
  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

@@

```{r}
# Which are the top n most frequently used words in each thesis?
get_top_words_per_document(data = emu_theses_words, 
                           top_n = 3)
```

```{r}
# Which are the top n most frequently used words in all theses?
get_top_words_per_corpus(data = emu_theses_words, 
                         top_n = 20)
```
